The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-06-20 manuscript.Rmd
[fix #20] mention extinction of marine inverts
[fix #21] restructure intro + talk more about DNA

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities is limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies and their size range, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016], making it difficult to monitor changes in species distributions and abundances. There is therefore a pressing need to improve our knowledge of marine systems before their functioning is irreversibly altered, and species go extinct.

@@

The proportion of species that have DNA barcodes available is relatively constant across the better studied phyla at the scale of the EEZ. It varies between 12% for bryozoans to 28% for cnidarians. The proportions for species known to occur in the Gulf of Mexico are usually similar than for the US, while typically a higher fraction of species from the Pacific Northwest are available. In 2010, it was estimated that a global scale, less than 10% of the species for each marine phylum had been sequenced [@Radulovici2010], and that this number was even lower in some areas (e.g. 2.6% for the Antarctic [@Grant2009]). This suggests that even as many additional species have had their DNA barcodes made available since these studies, the US fauna is generally better characterized than other parts of the world.

@@

@@

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems.
Digitization efforts are catching up on an already fragmented and vastly incomplete records.

>>>>

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities is limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies and their size range, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016], making it difficult to monitor changes in species distributions and abundances. While most documented global extinctions of marine species are vertebrates, many others may have been lost unnoticed because they were not known [@Sala2006]. There is therefore a pressing need to improve our knowledge of marine systems before their functioning is irreversibly altered, and species go extinct.

Research attempting to understand how much we know about marine biodiversity has focused on estimating the total number of species that inhabits the oceans, and which fraction of this diversity has been named [@Appeltans2012,@Costello2013]. Undoubtely, the development of molecular methods to investigate species limits has reinvigorated taxonomy, and contributed to elucidating cryptic species complexes [@Knowlton2000]. Thus, it has been argued that molecular data can speed up the enterprise of discovering and naming species [@Blaxter2016]. As species remain the most convenient unit to measure biological diversity, understanding how much has been done, how much remains to be done, and how we can speed up the description of species diversity is crucial. However, to understand the role of these species in their environment, we need to document the ecology of these species [@Cernansky2017]. For instance, to be able to predict how their geographic ranges will be affected by global change, we need detailed information about their environmental requirements. Occurrence data for marine species could be used for environmental niche modelling of their geographic and bathymetric distributions but the proportion of species for which we have enough data to do so is unknown.

@@

The proportion of species that have DNA barcodes available is relatively constant across the better studied phyla at the scale of the EEZ. It varies between 12% for bryozoans to 28% for cnidarians. The proportions for species known to occur in the Gulf of Mexico are usually similar than for the US, while typically a higher fraction of species from the Pacific Northwest are available. In 2012, it was estimated that a global scale, less than 3% of the marine species had been sequenced [@Appeltans2012], and that this number was even lower in some areas (e.g. 2.6% for the Antarctic [@Grant2009]). This suggests that even as many additional species have had their DNA barcodes made available since these studies, the US fauna is generally better characterized than other parts of the world.

@@

Considering additional linked data with species such as DNA has been a first step in a digital biodiversity world where other data on species can be gathered in databases and analyzed to better understand species distributions. This has paved the way to more linked data/databases. Still issues though with connecting identifiers [@Guralnick2014].

@@

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems.
Digitization efforts are catching up on an already fragmented and vastly incomplete records. Here, we only reported results from the US where there is at least two centuries of biodiversity exploration data and high-level of digitization. Many other parts of the world also need additional attention, especially in some developing countries where environmental pressure on ecosystems is even higher than in the US.
